Classical Planning Algorithms ¡ on the Atari Video Games: Preliminary Results

### 2014-11-18 icaps15-paper/paper.tex
Edits to Optimality and Duplicates Section
* Clarified the sentence where we observe that accounting for alternative
  paths to the same state with differring rewards, guarantees that 2BFS will
  be optimal if given enought time.
Table added to repo

\begin{table*}
\centering
% \addtolength{\tabcolsep}{-2pt}
% {\scriptsize
{\small
\begin{tabular}{@{}l|r|rr|rr|rr|rr@{}}
%% table header
% \hline\\[-3ex]
%\hline
\multicolumn{1}{c|}{} & \multicolumn{1}{c|}{Random} & \multicolumn{2}{c|}{\BRFS}& \multicolumn{2}{c|}{$\IW(1)$}& \multicolumn{2}{c|}{\BFS}&
   \multicolumn{2}{c}{\UCT}
% \\[.2ex]
\\
% \hline\\\\[-3ex]
\hline
%% table body
\multicolumn{1}{c|}{Game}  &
% \multicolumn{1}{c|}{Avg. Score} &
% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c}{Avg. Time}\\[.2ex]
 \multicolumn{1}{c|}{Score} &
 \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
 \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
 \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
 \multicolumn{1}{c}{Score} & \multicolumn{1}{c}{Time}
% \\[.2ex]
\\[.2ex]
% \hline\\
\hline
\Asterix        & 210    & 3900  & 32.24 & 119000  & 17.08 & 56320  & 0.95  & \textbf{296300} (290700) & 10.82\\
\BeamRider     & 448.8 & 519.2 & 43.48 & 3000          & 6.02  & 3804        & 0.92  & \textbf{4320} (6624)   & 42.91\\
\Breakout       & 1    & 2     & 51.56 & 383.6         & 1.37  & 421.2       & 3.95  & \textbf{512.4} (364)   & 26.03\\
\Enduro          & 0   & 0.4  & 40.92 & \textbf{500}   & 38.13 & 292.8      & 13.12 & 306.8 (286)             & 32.14\\
\Freeway         & 0   & 0     & 71.84 & \textbf{29.4} & 15.19 & 19.6       & 4.44  & 1 (0)    & 90.64\\
\Pong           & -20.4 & -20.2 & 41.51 & 16.4         & 18.45 & \textbf{21} & 39.69 & \textbf{21} (21) & 38.37\\
\Qbert           & 185   & 160  & 42.83 & 1545         & 1.82  & 14725       & 1.76  & \textbf{18730} (17343) & 28.61\\
\Seaquest       & 192   & 200   & 40.01 & \textbf{4274}& 14.72 & 4070        & 1.59  & 3406 (5132)      & 26.16\\
\SpaceInvaders & 187    & 115   & 46.11 & \textbf{2670} & 12.1  & 1327       & 2.25  & 2634 (2718)  & 19.66
\end{tabular}
}

 \caption{\small Performance that results from random policy and various  lookahead algorithms in 9 Atari games.
The algorithms  are \BRFS, $\IW(1)$, \BFS, and \UCT. Averages  over 5 runs shown with ``random'' seeds
$0$ to $4$, and maximum episode duration of $18,000$  frames. Every algorithm limited to a budget of
100,000 simulated frames, except for UCT limited to 500 rollouts, 300 frames deep instead. Average scores and CPU times per action selection shown.
Performance of UCT results  reported in \cite{bellemare:jair2013} shown in parenthesis.
Numbers  in bold show best performer in terms of average score. }
\label{performance}

@@

The notions of width and the \IW\  algorithm guarantee that  states with low width will be generated
in low polynomial time through \emph{shortest} paths. In the presence of rewards like the Atari games,
however, the  interest is not  in the shortest paths but in the \emph{best} paths; i.e, the paths with
maximum reward.  \IW\ may actually  fail to find such paths even when calling \IW($k$) with a high 
$k$ parameter. Optimality  could be  achieved then  by replacing the breadth-first search
underlying \IW($k$) by Dijkstra's algorithm yet such a move would make the relation between
\IW\ and the notion of width less clear. Another option is to change  \IW\ to comply with 
a different property; namely, that if there is a ``rewarding'' path made up of states of 
low width, then \IW\ will find such paths or better ones in time that is exponential 
in their width. For this actually a simple change in \IW\ suffices: when generating a state 
$s$ that is a \emph{duplicate} of a state $s'$ that has been previously generated and not pruned, 
$s'$ is replaced by $s$ if $R(s) > R(s')$. Moreover in such a case, the change in reward 
must be propagated to the descendants of $R(s')$ that are in memory. This is similar to the change required 
in A* for preserving optimality when moving from consistent to inconsistent heuristics \cite{pearl:heuristics}. 
The alternative is to ``reopen'' such nodes.  The same change is actually in \BFS\ to ensure that, 
if given enough time, \BFS\ will actually find the optimal paths. The code used for \IW\ and \BFS\ in the
experiments above does not implement this change, however, as the overhead involved in checking for duplicates 
in some test cases did not appear to pay off. More experimentation is probably needed.

>>>>

\begin{table*}
\centering
% \addtolength{\tabcolsep}{-2pt}
% {\scriptsize
{\small
\begin{tabular}{@{}l|rr|rr|r|r@{}}
%% table header
% \hline\\[-3ex]
%\hline
\multicolumn{1}{c|}{}& \multicolumn{2}{c|}{$\IW(1)$}& \multicolumn{2}{c|}{\BFS} & \multicolumn{1}{c|}{\BRFS} &  \multicolumn{1}{c}{\UCT}
% \\[.2ex]
\\
% \hline\\\\[-3ex]
\hline
%% table body
\multicolumn{1}{c|}{Game}  &
% \multicolumn{1}{c|}{Avg. Score} &
% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c}{Avg. Time}\\[.2ex]
 \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
 \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
 \multicolumn{1}{c|}{Score} &
 \multicolumn{1}{c}{Score} 
% \\[.2ex]
\\[.2ex]
% \hline\\
\hline

\propername{Assault} & 806.4 & 18.21 & \textbf{852.6} & 25.11 & 413.7 & 1512.2 \\ \hline
\propername{Asterix} & 153400 & 24.33 & 77200 & 26.58 & 2135.7 & \textbf{290700} \\ \hline
\propername{Asteroids} & \textbf{51338} & 65.92 & 22168 & 64.53 & 3127.4 & 4660.6 \\ \hline
\propername{Atlantis} & 159420 & 12.8 & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 30460 & \textbf{193858} \\ \hline
\propername{Bank\_heist} & \textbf{717} & 39.29 & 362 & 63.83 & 21.5 & 497.8 \\ \hline
\propername{Battle\_zone} & 11600 & 85.96 & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 6312.5 & \textbf{70333.3} \\ \hline
\propername{Beam\_rider} & 9108 & 22.53 & \textbf{9297.6} & 29.03 & 693.5 & 6624.6 \\ \hline
\propername{Berzerk} & \textbf{2096} & 58 & 802 & 73.47 & 195 & 553.5 \\ \hline
\propername{Bowling} & \textbf{68.6} & 9.84 & 50.4 & 59.52 & 25.5 & 25.1 \\ \hline
\propername{Boxing} & 100 & 15.2 & 100 & 22.3 & 100 & 100 \\ \hline
\propername{Breakout} & 384.2 & 4.22 & \textbf{771.6} & 39.47 & 1.1 & 364.4 \\ \hline
\propername{Centipede} & 99207.2 & 39.14 & 93067 & 65.49 & \textbf{125123} & 110422 \\ \hline
\propername{Chopper\_command} & 2100 & 41.06 & 3800 & 40.94 & 1827.3 & \textbf{34018.8} \\ \hline
\propername{Crazy\_climber} & 36160 & 3.61 & 36940 & 58.12 & 37110 & \textbf{98172.2} \\ \hline
\propername{Enduro} & \textbf{500} & 65.69 & 359.4 & 37.66 & 0.6 & 286.3 \\ \hline
\propername{Freeway} & \textbf{30.6} & 31.69 & 22.6 & 60.77 & 0 & 0.4 \\ \hline
\propername{Frostbite} & \textbf{3582} & 12.25 & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 137.2 & 270.5 \\ \hline
\propername{Gravitar} & 3920 & 61.85 & \textbf{5980} & 62.45 & 395 & 2850 \\ \hline
\propername{Hero} & \textbf{12887} & 36.63 & 7096.6 & 51.83 & 1323.8 & 12859.5 \\ \hline
\propername{Jamesbond} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & \textbf{8550} & 25.68 & 25 & 330 \\ \hline
\propername{Kangaroo} & \textbf{5700} & 7.9 & 5320 & 31 & 90 & 1990 \\ \hline
\propername{Krull} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & \textbf{5124} & 36.82 & 3089.2 & 5037 \\ \hline
\propername{Montezuma\_revenge} & 0 & 17.36 & \textbf{80} & 34.72 & 0 & 0 \\ \hline
\propername{Ms\_pacman} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 14781 & 23.97 & 1708.5 & \textbf{22336} \\ \hline
\propername{Pong} & \textbf{21} & 16.67 & \textbf{21} & 34.92 & -20.7 & \textbf{21} \\ \hline
\propername{Pooyan} & 11225 & 7.97 & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 909.7 & \textbf{17763.4} \\ \hline
\propername{Qbert} & 3705 & 11.07 & 11680 & 34.61 & 132.8 & \textbf{17343.4} \\ \hline
\propername{Riverraid} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & \textbf{5062} & 36.53 & 2178.5 & 4449 \\ \hline
\propername{Seaquest} & \textbf{14272} & 24.91 & 6138 & 33.38 & 288 & 5132.4 \\ \hline
\propername{Space\_invaders} & 2877 & 21.37 & \textbf{3974} & 34.21 & 112.2 & 2718 \\ \hline
\propername{Star\_gunner} & 1540 & 18.89 & \textbf{4660} & 17.62 & 1345 & 1207.1 \\ \hline
\propername{Tennis} & \textbf{23.8} & 21.05 & 23.6 & 36.33 & -23.8 & 2.8 \\ \hline
\propername{Venture} & 775 & 23.41 & \textbf{1080} & 27.78 & 0 & 0 \\ \hline
\propername{Video\_pinball} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 62074.6 & 43.13 & 55567.3 & \textbf{254748} \\ \hline
\propername{Zaxxon} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 15680 & 31.01 & 0 & \textbf{22610} \\ \hline
\multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\ \hline
Times Best (35 games)& \multicolumn{2}{c|}{12} & \multicolumn{2}{c|}{12} & 1 & 11 \\ \hline
Times Better than UCT (35 games) & \multicolumn{2}{c|}{23} & \multicolumn{2}{c|}{22} & 1  & -- \\ \hline

\end{tabular}
}

 \caption{\small Performance that results from random policy and various  lookahead algorithms in 35 Atari games.
The algorithms  are \BRFS, $\IW(1)$, \BFS, and \UCT. Averages  over 5 runs shown with ``random'' seeds
$0$ to $4$, and maximum episode duration of $18,000$  frames. Every algorithm limited to a budget of
150,000 simulated frames, except for UCT limited to 500 rollouts, 300 frames deep instead. Average scores and CPU times per action selection shown.
Performance of UCT taken from \cite{bellemare:jair2013}.
Numbers  in bold show best performer in terms of average score.
--Domains containing (--) are still computing,  each run can take 24 hours of computation. All domains from \cite{bellemare:jair2013}  will be ready for the camera ready--. }
\label{performance}

@@

The notions of width and the \IW\  algorithm guarantee that  states with low width will be generated
in low polynomial time through \emph{shortest} paths. In the presence of rewards like the Atari games,
however, the  interest is not  in the shortest paths but in the \emph{best} paths; i.e, the paths with
maximum reward.  \IW\ may actually  fail to find such paths even when calling \IW($k$) with a high 
$k$ parameter. Optimality  could be  achieved then  by replacing the breadth-first search
underlying \IW($k$) by Dijkstra's algorithm yet such a move would make the relation between
\IW\ and the notion of width less clear. Another option is to change  \IW\ to comply with 
a different property; namely, that if there is a ``rewarding'' path made up of states of 
low width, then \IW\ will find such paths or better ones in time that is exponential 
in their width. For this actually a simple change in \IW\ suffices: when generating a state 
$s$ that is a \emph{duplicate} of a state $s'$ that has been previously generated and not pruned, 
$s'$ is replaced by $s$ if $R(s) > R(s')$. Moreover in such a case, the change in reward 
must be propagated to the descendants of $R(s')$ that are in memory. This is similar to the change required 
in A* for preserving optimality when moving from consistent to inconsistent heuristics \cite{pearl:heuristics}. 
The alternative is to ``reopen'' such nodes. 
%The same change is actually in \BFS\ to ensure that, 
%if given enough time, \BFS\ will actually find the optimal paths.
Changing \BFS\ to account for such paths guarantees that, if given enough time, \BFS\ will actually find
the optimal paths. 
The code used for \IW\ and \BFS\ in the
experiments above does not implement this change, however, as the overhead involved in checking for duplicates 
in some test cases did not appear to pay off. More experimentation is probably needed.
